NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-03-31 paper/appendix_modelDescription.tex
minor fixes

\section{Appendix B: Code Example: Link Visualization with Model}
From the user point of view, the following python script is all the code the domain expert needs to link the NLP model with the visual interface. The visualization library's API defines several callbacks (call Hook in the code) that describe the model functionality (e.g., prediction) need to be supplied by the NLP model.

>>>>

\section{Appendix B: Code Example: Link Visualization with Model}
From the user's point of view, the following Python script is all the code the domain expert needs to link the NLP model with the visual interface and generate the tool used in the paper. The visualization library's API defines several callbacks (called Hook in the code) that describe the model functionality (e.g., prediction) need to be supplied by the NLP model.

### 2018-03-31 paper/attention.tex
minor fixes

The matrix attention view (Fig.~\ref{fig:attentionVis}(b)), despite being more verbose and less efficient in highlighting the dominant alignment, does not have similar shortcomings. Together, the graph and matrix views complement each other and provide the same information from different perspectives. As illustrated by the pink arrow line, we can see how the same attention value is visualized in both the graph and the matrix view.
To help the user recognize the correspondence during the exploration, we enable the linkage between highlighted actions in both views.

To support the ability to perturb the attention values (\textbf{T2}), we include the attention editing functionality. The attention matrix view is the most suitable place to conduct the edit operation since it provides a direct mapping of attention value.
As we can see in Fig.~\ref{fig:attentionVis}(c), when a user clicks the cell of the matrix, a sider will pop up for customizing the attention value (as the user edits the value, each row is automatically normalized).
%
As illustrated in Fig.~\ref{fig:attentionVis}($a_{2}$)(f), we allow the user to compare currently and previously displayed attentions by computing and visualizing their difference.

>>>>

The matrix attention view (Fig.~\ref{fig:attentionVis}(b)), despite being more verbose and less efficient in highlighting the dominant alignment, does not have similar shortcomings. Together, the graph and matrix views complement each other and provide the same information from different perspectives. As illustrated by the pink arrowed lines, we can see how the same attention value is visualized in both the graph and the matrix view.
To help the user recognize the correspondence during the exploration, we enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), the attention of the two ``couple'' is highlighted).

To support the ability to perturb the attention values (\textbf{T2}), we include the attention editing functionality. The attention matrix view is the most suitable place to conduct the edit operation since it provides a direct mapping of attention value.
As we can see in Fig.~\ref{fig:attentionVis}(c), when a user clicks the cell of the matrix, a sider will pop up for customizing the attention value (as the user edits the value, each row is automatically renormalized).
%
As illustrated in Fig.~\ref{fig:attentionVis}($a_{2}$)(f), we allow the user to compare currently and previously displayed attentions by computing and visualizing their difference.

### 2018-03-31 paper/pipeline.tex
adjust figures
minor fixes

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{pipelineView}
 \vspace{-3mm}
 \caption{
The pipeline view provides the visual representation of three-stage (encoder, attention, classifier) model. 
%The pipeline view, by showing the aggregated gradient distribution, informs the user about how different parts of the model respond to the optimization.
 %
In the proposed tool, we allow the model parameter to be updated to correct a prediction error. 
The optimization hyperparameters are shown in (a), and each of the stages is illustrated in (b), where the user can click on the blue ``parameter'' area to enable or disable its update (the legend is shown in (d)). In (c), we select whether we want to use the current update setting as displayed or try all the configuration combinations as the parameter update can be enabled or disabled for each stage.
 }
\label{fig:pipelineView}
\vspace{-2mm}
\end{figure}

\subsection{Pipeline View}
\label{sec:pipeline}
The Pipeline view provides a direct visual representation of the three stages (encoder, attention, classifier) of the model. In the proposed tool, we allow the model parameter to be updated (via an optimization) to correct a prediction error (\textbf{T3}). The pipeline view, by visualizing the aggregated gradient distribution, informs the user about how each stage responds to the optimization.

There are many ways to update the model to correct a prediction. The simplest approach is applying standard backpropagation and overfits to the example. However, without any constraint, the update step may alter the model in unexpected ways.
% \shusen{what is the benefit of using mira?}
Instead, we adopt the idea in \emph{margin-infused relaxed algorithm (MIRA)}~\cite{CrammerSinger2003}, where we regulate the optimization with the L2-norm of the parameter change. Therefore, in the proposed tool, we obtain the target parameters by the following:
\begin{equation}
\underset{\mathbf{W}'}{\mathrm{argmin}}( C \mathbb{J}(\mathbf{W}') + ||\mathbf{W}' - \mathbf{W_o}||^2)
\end{equation}
where $\mathbb{J}(\mathbf{W})$ is the loss function of the neural model, $\mathbf{W_o}$ is the original model's parameters taken as constant, $\mathbf{W}'$ is the parameters to update, and $C$ is the weighting term, which determines whether we intended to emphasize more on obtaining better fitting or deviating less from the original model. Due to the nonconvex nature of the neural networks, we use SGD to optimize the above combined loss function.
%
With this formulation, we try to find a good approximation to the newly assigned label, while still maintaining relatively small change with respect to the original model.

The optimization hyperparameters are shown in Fig.~\ref{fig:pipelineView}(a). Each of the stages is illustrated by a glyph (Fig.~\ref{fig:pipelineView}(b)), in which the user can enable or disable its parameter update by clicking on the parameter bar (the legend about its state is shown in Fig.~\ref{fig:pipelineView}(d) ). In Fig.~\ref{fig:pipelineView}(c), we select whether we want to use the current update setting as displayed or try all the configuration combinations (i.e., each stage can either be enabled or disabled for the three stages; therefore, there are $8$ combinations in total, or $7$ we discard the case where all stages are not updated).

>>>>

\subsection{Pipeline View}
\label{sec:pipeline}
The Pipeline view provides a direct visual representation of the three stages (encoder, attention, classifier) of the model. In the proposed tool, we allow model parameters to be updated (via an optimization) to correct a prediction error (\textbf{T3}). The pipeline view, by visualizing the distribution of the parameter changes, informs the user about how each stage responds to the optimization.

There are many ways to update the model to correct a prediction. The simplest approach is applying standard backpropagation and overfits to the example. However, without any constraint, the update step may alter the model in unexpected ways.
% \shusen{what is the benefit of using mira?}
Instead, we adopt the idea in \emph{margin-infused relaxed algorithm (MIRA)}~\cite{CrammerSinger2003}, where we regulate the optimization with the L2-norm of the parameter change. In the proposed tool, we obtain the target parameters by the following optimization:
\begin{equation}
\underset{\mathbf{W}'}{\mathrm{argmin}}( C \mathbb{J}(\mathbf{W}') + ||\mathbf{W}' - \mathbf{W_o}||^2)
\end{equation}
where $\mathbb{J}(\mathbf{W})$ is the loss function of the neural network model, $\mathbf{W_o}$ is the original model's parameters taken as constant, $\mathbf{W}'$ is the updated parameters, and $C$ is the weighting term, which determines whether we intended to emphasize more on obtaining better fitting or deviating less from the original model. Due to the nonconvex nature of the neural networks, we use SGD to optimize the above combined loss function.
%
With this formulation, we try to find a good approximation to the newly assigned label, while still maintaining relatively small change with respect to the original model.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{pipelineView}
 \vspace{-2mm}
 \caption{
The pipeline view provides the visual representation of the three-stage (encoder, attention, classifier) model. 
%
In the proposed tool, we allow model parameters to be updated to correct a prediction error. 
The optimization hyperparameters are shown in (a), and each of the stages is illustrated in (b), where the user can click on the blue rectangle marked with ``parameter'' to enable or disable its update (the legend is shown in (d)). In (c), we select whether we want to use the current pipeline setting as displayed or try all the pipeline configuration combinations as the parameter update operation can be enabled or disabled for each stage.
 }
\label{fig:pipelineView}
\vspace{-2mm}
\end{figure}

The optimization hyperparameters are shown in Fig.~\ref{fig:pipelineView}(a). Each of the stages is illustrated by a glyph (Fig.~\ref{fig:pipelineView}(b)), in which the user can enable or disable its parameter update by clicking on the blue rectangle marked with ``parameter'' (the legend about its state is shown in Fig.~\ref{fig:pipelineView}(d) ). In Fig.~\ref{fig:pipelineView}(c), we select whether we want to use the current update setting as displayed or try all the configuration combinations (i.e., each stage can either be enabled or disabled for the three stages; therefore, there are $8$ combinations in total, or $7$ we discard the case where all stages are not updated).

### 2018-03-31 paper/prediction.tex
space squeeze
update task
adjust figures
minor fixes

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradiction, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}(a), a prediction probability is encoded as a point in the barycentric coordinate system of the triangle. 
Let $\mathbf{C_1}$,$\mathbf{C_2}$,$\mathbf{C_3}$ be the three points defines a triangle, and let $p_1$, $p_2$, $p_3$ be the probabilities the label is neutral, contradiction and entailment, respectively. The coordinate of the prediction $\mathbf{C_p}$ in the triangle is computed as $\mathbf{C_p} = p_1\mathbf{C_1}+p_2\mathbf{C_2}+p_3\mathbf{C_3}$.
In the triangle, the distinctly colored background indicates the regions that correspond to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas the smaller gray circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.

To perturb the prediction (\textbf{T3}), we need a way to communicate the reassignment of the predicted label. As illustrated in Fig.~\ref{fig:predictionView}(b), we integrate such an operation in the prediction view. When pressing and dragging the prediction (represented as a circle), the user is presented with the three options (E, N, C) corresponding to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label. The assignment is applied when the user releases the mouse while hovering on the label of choice.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-2mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
 %
% The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth. Therefore, we apply a label reassignment operation, which triggers a model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
 }
\label{fig:predictionView}
\end{figure}

>>>>

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradiction, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}(a), a prediction probability is encoded as a point in the barycentric coordinate system of the triangle. 
Let $\mathbf{C_1}$,$\mathbf{C_2}$,$\mathbf{C_3}$ be the three points defines a triangle, and let $p_1$, $p_2$, $p_3$ be the probabilities the label is neutral, contradiction or entailment. The coordinate of the prediction $\mathbf{C_p}$ in the triangle is computed as $\mathbf{C_p} = p_1\mathbf{C_1}+p_2\mathbf{C_2}+p_3\mathbf{C_3}$.
In the triangle, the distinctly colored background (grey, light grey, and light green) indicates the regions that correspond to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas the smaller gray circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-5mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
 %
% The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth. Therefore, we apply a label reassignment operation, which triggers a model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
 }
  \vspace{-2mm}
\label{fig:predictionView}
\end{figure}

To perturb the prediction (\textbf{T3}), we need a way to communicate the reassignment of the predicted label. As illustrated in Fig.~\ref{fig:predictionView}(b), we integrate such an operation in the prediction view. When pressing and dragging the prediction (represented as a circle), the user is presented with the three options (E, N, C) corresponding to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label. The reassignment is applied when the user releases the mouse while hovering on the label of choice.

### 2018-03-31 paper/sentence.tex
add open source comment add title to the appendix
minor fixes

\subsection{Sentence View}
\label{sec:sentence}
% Once an exploration cycle started, the experts need to examine the premise and hypothesis sentence the current example.
As illustrated in Fig.~\ref{fig:teaser}(B), the sentence view shows the premise and hypothesis sentence.
%
To facilitate the analysis task (\textbf{T1}), we employ an automate sentence perturbation scheme that replaces \emph{nouns} and \emph{verbs} by their synonyms in the wordNet~\cite{Miller1995} (the standard lexical database for NLP applications).
%
When the perturbation is applied to either of the sentences, the replaced words are highlighted in blue to signal the modification made to the original sentence.

@@

At the top left of the view, the two controls \emph{Predict} and \emph{Predict All} correspond to predicting the currently displayed sentence pairs and predict all combinations of perturbed premises and hypotheses, respectively.
%
To avoid the situation where both sentences are perturbed, we ensure only one perturbation per sentence pair (i.e., we use the original premise if the hypothesis is perturbed, or use the original hypothesis if the premise is perturbed).
%%% sentence perturbation %%%%
The previously explored original sentences are stored in the dropdown list (left of the buttons) that allow the user to revisit examined examples.
Also, the user can type any sentences or modified existing text in the sentence display areas to accommodate user-defined input.

>>>>

\subsection{Sentence View}
\label{sec:sentence}
% Once an exploration cycle started, the experts need to examine the premise and hypothesis sentence the current example.
As illustrated in Fig.~\ref{fig:teaser}(b), the sentence view shows the premise and hypothesis sentence.
%
To facilitate the analysis task (\textbf{T1}), we employ an automate sentence perturbation scheme that replaces \emph{nouns} and \emph{verbs} by their synonyms in the wordNet~\cite{Miller1995} (the standard lexical database for NLP applications).
%
When the perturbation is applied to either of the sentences, the replaced words are highlighted in blue to signal the modification made to the original sentence.

@@

At the top left of the view, the two controls \emph{Predict} and \emph{Predict All} correspond to predicting the currently displayed sentence pairs and predict all combinations of perturbed premises and hypotheses, respectively.
%
To avoid the situation where both sentences are perturbed, we ensure only one perturbation per sentence pair (i.e., we use the original premise if the hypothesis is perturbed, or use the original hypothesis if the premise is perturbed).
%%% sentence perturbation %%%%
The previously explored original sentences are stored in the dropdown list (on the right side of the buttons) that allow the user to revisit examined examples.
Also, the user can type any sentences or modify existing text in the sentence display areas to accommodate user-defined inputs.
